Subsegmental language detection in Celtic language text

### 2014-07-02 codeswitch-cltw-2014/codeswitch_camera_ready.tex
maybe there
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55076 72bbbca6-d526-0410-a7d9-f06f51895060
FINAL
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55078 72bbbca6-d526-0410-a7d9-f06f51895060
comments from teresa
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55176 72bbbca6-d526-0410-a7d9-f06f51895060

\section{Introduction}
\label{intro}
Determining the language of a piece of text is one of the first steps that must be taken
before proceeding with further computational processing. This task has received a substantial amount of
attention in recent years \cite{cavnar1994n,lui2012langid}. However, previous research has on the whole assumed
that a given text will be in a single language. When dealing with text from formal domains,
this may be the case --- although there are exceptions --- such as quotations embedded in
the text in another language. But when dealing with informal text, particularly in languages
where the speech community is predominantly bi- or multi-lingual, this assumption may not hold.
%Such as
%code switching, quotations, named entities, interjections, translations --- either of the full text, or 
%part of the text, comments, etc.

@@

The remainder of the paper is laid out as follows. In section~\ref{sec:method} we describe the problem
in more detail and look at relevant prior work before proposing a novel method of sub-sentential
language detection. Section~\ref{sec:eval} describes the evaluation methodology. Then in section~\ref{sec:results}
we present the results of our method and compare it against several other possible methods. Finally, section~\ref{sec:conclusions}
presents future work and conclusions.
\blfootnote{
    %
    % for review submission
    %
    %\hspace{-0.65cm}  % space normally used by the marker
    %Place licence statement here for the camera-ready version, see
    %Section~\ref{licence} of the instructions for preparing a
    %manuscript.
    %
     % final paper: en-uk version (to license, a licence)

@@

\section{Related Work}
\label{sec:relwork}

@@

\section{Methodology}
\label{sec:method}

@@

\subsection{Corpus}

@@

\begin{figure}
\begin{small}
\begin{alltt}
[\textbf{en} You're a] [\textbf{ga} Meirice\'{a}nach, c\'{e}n f\'{a}th] [\textbf{en} are you] [\textbf{ga} foghlaim Gaeilge?!] 
@afaltomkins [\textbf{cy} gorfod cael bach o tan] [\textbf{en} though init] 
[\textbf{en} omg] [\textbf{cy} mar cwn bach yn] [\textbf{en} black and tan] [\textbf{cy} a popeth,] [\textbf{en} even cuter!!] 
\end{alltt}
\end{small}
\vspace{-0.3cm}
\caption{Example of text from a microblogging site chunked manually.}
\label{fig:tweets}
\end{figure}

\subsection{Alphabet n-gram approach}
\label{alphan}

@@

\subsection{Sequence chunking}
\label{cschunking}

This section describes the way we apply heuristics to segment and label the input text.
%the heuristics taken into account while performing the chunking of the input data according to 
%the sequences of segments in different languages. 
In Figure ~\ref{fig:chunkal}, `chunks' represent the list of evaluated tuples of segments and their labelled language, `buffer' is the expanding segment.
 \textit{lang\_predict} corressponds to any function which is used to determine the language of the token. The flag 
variable helps in implementing the heuristic of minimum segment size while labelling chunks.
\begin{algorithm}
\caption*{\textsc{}}
%\caption{Chunking algorithm}\label{euclid}
{\fontsize{9}{9}\selectfont
\begin{algorithmic}[1]
%\Statex{}%Procedure{}{}
\State $\textit{buffer} = [\ ] \text{\ \ \  \#undecided expanding window of chunk}$
\State $\textit{chunks} = [\ ] \text{\ \ \  \#decided labelled segment}$
\State $\textit{buffer\_language} \gets \textit{lang\_predict}\text{(sentence[0])} \text{\ \ \ \#Language of first word}$
\State $\textit{flag} \gets \text{0}$
%\State $\text{for \ } \textit{word \ } \text{in \ } \textit{sentence:} $
\For{\texttt{word in sentence}}
\If {$ \textit{lang\_predict(word)} \text{==} \textit{buffer\_language} $}
        \If {$ \textit{flag} \text{=\ =\ 1} $}
        \State $\textit{buffer} += [\textit{word\_d,word}] $
        \State $\textit{flag} \gets \text{0}$
        \Else
        \State $\textit{buffer} += [\textit{word}] $
        \EndIf
        \EndIf
\If {$ \textit{lang\_prediction(word)} \text{!=} \textit{buffer\_langugae} $}
        \If {$ \textit{flag} \text{=\ =\ 0} $}
        \State $\textit{flag} \gets \text{1}$
        \State $\textit{word\_d} \gets \textit{word} $
        \State $\text{continue}$

@@

\EndIf
\EndFor
\If {$\textit{length(buffer)} \text{!=0\ } $}
        \State $\textit{chunk} \text{+= \ }  \text{[\textit{[(buffer,buffer\_language)]}} $

\EndIf
%\EndProcedure
%\Statex{}
\end{algorithmic}}
\end{algorithm}
\begin{figure}[hb]
\def\@fs@post{}
\caption{Chunking Algorithm}
\label{fig:chunkal}
\end{figure}
%A flag is set at the beginning of the input text, and this moves forward to the position in the text upto where confident 
%chunking has been performed. For example, in case of a continuous prediction there would be no change and since the 
%predicted tag of the new word corresponds to that already of the expanding chunk, this word is also included in the 
%same chunk. In case of a doubt, the flag is set at the same position, and no final decision on labelling the word is 
%performed, to have more confidence on the code-switch, the process moves on to the next token, if it corresponds to 
%that of the previous chunk, then both these undecided words are labelled and included in the chunk, otherwise a change 
%is noted and a new chunk is created with the changed label. In this process, we are trying to identify the sub-segments 
%that we need, thus performing both tasks at once, the first being language prediction and second being the chunking decision.

\vspace{-0.7cm}
\subsection{Word-based prediction}
\vspace{-0.132cm}
\label{wbased}

@@

%A better way to predict the spans of the sub-segments of the text is to include the two methods of word and character-based techniques 
%as described above.
In case of the word being present in only one of the two monolingual word lists the classification is simple, 
but in case of a conflict, a character bigram backoff was introduced to help us disambiguate the language label. This method works well because the earlier heuristic approach of just labelling the word with the label of the span which is expanding, would mean less code-switch detection and more shift towards the majority class.

@@

\section{Evaluation}
\vspace{-0.152cm}
\label{sec:eval}

@@

Apart from the techniques discussed in section~\ref{sec:method}, some baselines are also used to give a comparative view of how well all the mechanisms perform.

\subsection{Baseline}
\label{baseline}

@@

\subsection{\texttt{langid} character trigram prediction}
\label{langidstuff}

For this system we used the character trigram probabilities to predict the detected language for each token. Trigrams were chosen after experimenting with 1--5 grams. The heuristics in section~\ref{sec:method} were followed for the text processing and chunking part of the method.

@@

\begin{table}
\begin{center}
\begin{tabular}{|l|r|r|r|}
\hline
\multirow{2}{*}{\textbf{System}} &  \multicolumn{3}{c|}{\textbf{Accuracy} (\%)} \\\cline{2-4}
       &   Irish---English & Welsh---English & Breton---French \\ 
\hline
\texttt{baseline} & 42.76 & 42.16 & 44.07 \\
\hline
\texttt{langid-3character} & 57.24 & 45.92 & 43.16 \\
\hline
\texttt{wordlist} & 79.75 & \textbf{74.28} & 83.96 \\
\hline
\texttt{character bigram} & 81.29 & 65.62 & 76.79 \\
\hline
\texttt{wordlist+character bigram} & \textbf{85.79} & 72.40 & \textbf{88.79} \\
\hline
\end{tabular}
\end{center}
\vspace{-0.3cm}
\caption{Accuracy of the systems over the three language pairs. The accuracy measures how often a token
  was assigned to the right language, independent of span.}
\label{table:accuracy}
\end{table}
%In terms of per-word accuracy, the baseline performs as expected, given that half of the words in the test set are from a given language. 
%The \texttt{langid.py} based methods allow for segmenting the text into chunks, but favour the majority language.
%\footnote{This could be because of the default models being used.} 
While the precision and recall are low for the remaining models, we see that we are able to improve 
The performance by combining the wordlist based model with a character bigram model. And what is more, we are able to begin to not only 
identify particular words in a language, but also segments.

\section{Conclusions}
\vspace{-0.152cm}
\label{sec:conclusions}

This paper has presented a very preliminary investigation into subsegment language identification in Celtic language texts. We have
proposed a model that chunks input text into segments and performs language identification on these segments at the same time. Precision and
recall are low, leaving a lot of room for further work.
% Some of our plans are as follows. We would like to annotate more test data,
%at least 100 tweets in each language.
 Although In \cite{king2013labeling}, the authors label on a per word level, yet we would like to include supervised methods and features talked about in this research to improve our efficiency while dealing with segments. We would also like to attempt our method using higher order character n-gram models for backoff, and n-gram word language models for detection and on more annotated data.

@@

\section*{Acknowledgements}

>>>>

\section{Introduction}
\vspace{-0.132cm}
\label{intro}
Determining the language of a piece of text is one of the first steps that must be taken
before proceeding with further computational processing. This task has received a substantial amount of
attention in recent years \cite{cavnar1994n,lui2012langid}. However, previous research has on the whole assumed
that a given text will be in a single language. When dealing with text from formal domains,
this may be the case --- although there are exceptions --- such as quotations embedded in
the text in another language. But when dealing with informal text, particularly in languages
where the speech community is predominantly bi- or multi-lingual, this assumption may not hold.
%Such as
%code switching, quotations, named entities, interjections, translations --- either of the full text, or 
%part of the text, comments, etc.

@@

The remainder of the paper is laid out as follows. In Section~\ref{sec:method} we describe the problem
in more detail and look at relevant prior work before proposing a novel method of sub-sentential
language detection. Section~\ref{sec:eval} describes the evaluation methodology. Then in Section~\ref{sec:results}
we present the results of our method and compare it against several other possible methods. Finally, Section~\ref{sec:conclusions}
presents future work and conclusions.
\blfootnote{
    %
    % for review submission
    %
    %\hspace{-0.65cm}  % space normally used by the marker
    %Place licence statement here for the camera-ready version, see
    %Section~\ref{licence} of the instructions for preparing a
    %manuscript.
    %
     % final paper: en-uk version (to license, a licence)

@@

\section{Related Work}
\vspace{-0.132cm}
\label{sec:relwork}

@@

\section{Methodology}
\vspace{-0.132cm}
\label{sec:method}

@@

\subsection{Corpus}
\vspace{-0.152cm}

@@

\begin{figure}
\begin{small}
\begin{alltt}
[\textbf{en} You're a] [\textbf{ga} Meirice\'{a}nach, c\'{e}n f\'{a}th] [\textbf{en} are you] [\textbf{ga} foghlaim Gaeilge?!] 
@afaltomkins [\textbf{cy} gorfod cael bach o tan] [\textbf{en} though init] 
[\textbf{en} omg] [\textbf{cy} mar cwn bach yn] [\textbf{en} black and tan] [\textbf{cy} a popeth,] [\textbf{en} even cuter!!] 
\end{alltt}
\end{small}
\caption{Example of text from a microblogging site chunked manually.}
\label{fig:tweets}
\end{figure}

\subsection{Alphabet n-gram approach}
\vspace{-0.152cm}
\label{alphan}

@@

\subsection{Sequence chunking}
\vspace{-0.152cm}
\label{cschunking}

This section describes the way we apply heuristics to segment and label the input text.
%the heuristics taken into account while performing the chunking of the input data according to 
%the sequences of segments in different languages. 
In Figure ~\ref{fig:chunkal}, `chunks' represent the list of evaluated tuples of segments and their labelled language, `buffer' is the expanding segment.
 \textit{lang\_predict} corresponds to any function which is used to determine the language of the token. The flag 
variable helps in implementing the heuristic of minimum segment size while labelling chunks.
\begin{algorithm}
\caption*{\textsc{}}
%\caption{Chunking algorithm}\label{euclid}
{\fontsize{9}{9}\selectfont
\begin{algorithmic}[1]
%\Statex{}%Procedure{}{}
\State $\textit{buffer} = [\ ] \text{\ \ \  \#Undecided expanding window of chunk}$
\State $\textit{chunks} = [\ ] \text{\ \ \  \#Decided labelled segment}$
\State $\textit{buffer\_language} \gets \textit{lang\_predict}\text{(sentence[0])} \text{\ \ \ \#Language of first word}$
\State $\textit{flag} \gets \text{0}$
%\State $\text{for \ } \textit{word \ } \text{in \ } \textit{sentence:} $
\For{\texttt{word in sentence}}
\If {$ \textit{lang\_predict(word)} \text{==} \textit{buffer\_language} $}
        \If {$ \textit{flag} \text{=\ =\ 1} $}
        \State $\textit{buffer} += [\textit{word\_d,word}] $
        \State $\textit{flag} \gets \text{0}$
        \Else
        \State $\textit{buffer} += [\textit{word}] $
        \EndIf
        \EndIf
\If {$ \textit{lang\_prediction(word)} \text{!=} \textit{buffer\_language} $}
        \If {$ \textit{flag} \text{=\ =\ 0} $}
        \State $\textit{flag} \gets \text{1}$
        \State $\textit{word\_d} \gets \textit{word} $
        \State $\text{continue}$

@@

\EndIf
\EndFor
\If {$\textit{length(buffer)} \text{!=0\ } $}
        \State $\textit{chunk} \text{+= \ }  \text{[\textit{(buffer,buffer\_language)}]} $

\EndIf
%\EndProcedure
%\Statex{}
\end{algorithmic}}
\end{algorithm}
\vspace{-0.4cm}
\begin{figure}[hb]
\def\@fs@post{}
\caption{Chunking Algorithm}
\label{fig:chunkal}
\end{figure}
\vspace{-0.4cm}
%A flag is set at the beginning of the input text, and this moves forward to the position in the text upto where confident 
%chunking has been performed. For example, in case of a continuous prediction there would be no change and since the 
%predicted tag of the new word corresponds to that already of the expanding chunk, this word is also included in the 
%same chunk. In case of a doubt, the flag is set at the same position, and no final decision on labelling the word is 
%performed, to have more confidence on the code-switch, the process moves on to the next token, if it corresponds to 
%that of the previous chunk, then both these undecided words are labelled and included in the chunk, otherwise a change 
%is noted and a new chunk is created with the changed label. In this process, we are trying to identify the sub-segments 
%that we need, thus performing both tasks at once, the first being language prediction and second being the chunking decision.

\subsection{Word-based prediction}
\vspace{-0.152cm}
\label{wbased}

@@

%A better way to predict the spans of the sub-segments of the text is to include the two methods of word and character-based techniques 
%as described above.
In case of the word being present in only one of the two monolingual word lists the classification is simple, 
but in case of a conflict, a character bigram backoff was introduced to help us disambiguate the language label. 
%This method works well because the earlier heuristic approach of just labelling the word with the label of the span which is expanding, would mean less code-switch detection and more shift towards the majority class.

@@

\vspace{-0.132cm}
\section{Evaluation}
\vspace{-0.132cm}
\label{sec:eval}

@@

Apart from the techniques discussed in Section~\ref{sec:method}, some baselines are also used to give a comparative view of how well all the mechanisms perform.

\subsection{Baseline}
\vspace{-0.132cm}
\label{baseline}

@@

\subsection{\texttt{langid} character trigram prediction}
\vspace{-0.132cm}
\label{langidstuff}

For this system we used the character trigram probabilities to predict the detected language for each token. Trigrams were chosen after experimenting with 1--5 grams. The heuristics in Section~\ref{sec:method} were followed for the text processing and chunking part of the method.

@@

\begin{table}
\begin{center}
\begin{tabular}{|l|r|r|r|}
\hline
\multirow{2}{*}{\textbf{System}} &  \multicolumn{3}{c|}{\textbf{Accuracy} (\%)} \\\cline{2-4}
       &   Irish---English & Welsh---English & Breton---French \\ 
\hline
\texttt{baseline} & 42.76 & 42.16 & 44.07 \\
\hline
\texttt{langid-3character} & 57.24 & 45.92 & 43.16 \\
\hline
\texttt{wordlist} & 79.75 & \textbf{74.28} & 83.96 \\
\hline
\texttt{character bigram} & 81.29 & 65.62 & 76.79 \\
\hline
\texttt{wordlist+character bigram} & \textbf{85.79} & 72.40 & \textbf{88.79} \\
\hline
\end{tabular}
\end{center}
\caption{Accuracy of the systems over the three language pairs. The accuracy measures how often a token
  was assigned to the right language, independent of span.}
\label{table:accuracy}
\end{table}
%In terms of per-word accuracy, the baseline performs as expected, given that half of the words in the test set are from a given language. 
%The \texttt{langid.py} based methods allow for segmenting the text into chunks, but favour the majority language.
%\footnote{This could be because of the default models being used.} 
While the precision and recall are low for the remaining models, we see that we are able to improve 
The performance by combining the wordlist based model with a character bigram model. And what is more, we are able to begin to not only 
identify particular words in a language, but also segments.

\vspace{-0.132cm}
\section{Conclusions}
\vspace{-0.132cm}
\label{sec:conclusions}

This paper has presented a very preliminary investigation into subsegment language identification in Celtic language texts. We have
proposed a model that chunks input text into segments and performs language identification on these segments at the same time. Precision and
recall are low, leaving a lot of room for further work.
% Some of our plans are as follows. We would like to annotate more test data,
%at least 100 tweets in each language.
 Although \newcite{king2013labeling} label on a per word level, yet we would like to include supervised methods and features talked about in this research to improve our efficiency while dealing with segments. We would also like to attempt our method using higher order character n-gram models for backoff, and n-gram word language models for detection and on more annotated data.

@@

\section*{Acknowledgements}
\vspace{-0.132cm}
